\bf Range Analysis for the IonMonkey JavaScript Compiler

### 2012-05-03 project-writings/poster.tex
Finish up the slides

\begin{slide}
\begin{center}
{\Huge \bf{Range Analysis}}
\end{center}
\begin{itemize}
\item Associate each integer temporary with a range: a contiguous subset of $\{-\infty\} \cup [-2^{31}, 2^{31}-1] \cup \{\infty\}$
\item Can do arithmetic on ranges:
\end{itemize}
\begin{eqnarray*}
\Clamp(n) &=& \begin{cases}
-\infty &\text{if } n < -2^{31} \\
\infty &\text{if } n > -2^{31}-1 \\
n&\text{otherwise } \\
\end{cases}\\
\Clamp([l, h]) &=& [\Clamp(l), \Clamp(h)] \\
{[x_l, x_h] \cup [y_l, y_h]} &=& [\min(x_l, y_l), \max(x_h, y_h)] \\
{[x_l, x_h] \cap [y_l, y_h]} &=& [\max(x_l, y_l), \min(x_h, y_h)] \\
{[x_l, x_h] + [y_l, y_h]} &=& \Clamp ([x_l + y_l, x_h + y_h]) \\
{[x_l, x_h] - [y_l, y_h]} &=& \Clamp ([x_l - y_h, x_h - y_l])
\end{eqnarray*}
\end{slide}

@@

>>>>

\begin{slide}
\begin{center}
{\Huge \bf{Range Analysis}}
\end{center}
\begin{itemize}
\item Associate each integer temporary with a range: a contiguous subset of $\{-\infty\} \cup [-2^{31}, 2^{31}-1] \cup \{\infty\}$
\item Infinite ranges indicate the operation can overflow; arithmetic operations with a finite range don't need overflow checks!
\item Can do arithmetic on ranges:
\end{itemize}
\begin{eqnarray*}
\Clamp(n) &=& \begin{cases}
-\infty &\text{if } n < -2^{31} \\
\infty &\text{if } n > -2^{31}-1 \\
n&\text{otherwise } \\
\end{cases}\\
\Clamp([l, h]) &=& [\Clamp(l), \Clamp(h)] \\
{[x_l, x_h] \cup [y_l, y_h]} &=& [\min(x_l, y_l), \max(x_h, y_h)] \\
{[x_l, x_h] \cap [y_l, y_h]} &=& [\max(x_l, y_l), \min(x_h, y_h)] \\
{[x_l, x_h] + [y_l, y_h]} &=& \Clamp ([x_l + y_l, x_h + y_h]) \\
{[x_l, x_h] - [y_l, y_h]} &=& \Clamp ([x_l - y_h, x_h - y_l])
\end{eqnarray*}
\end{slide}

@@

\begin{slide}
\begin{center}
{\Huge \bf{Range Analysis Algorithm}}
\end{center}
\begin{itemize}
\item Range analysis is simple and worklist based. Range recomputation is done with range arithmetic. $\beta$ intersects a range, $\phi$ unions.
\item Key insight: can always squash inputs to a finite range, since they would have bailed out if they overflowed.
\end{itemize}
{\tiny
\begin{verbatim}
def propagateRanges(graph):
    worklist = []
    for block in reversePostOrder(graph):
        for insn in block:
            insn.recomputeRange()
            if !insn.isPhi() and !insn.isBeta():
                continue
            worklist.appendAll(insn.uses())

    while len(worklist) > 0:
        insn = worklist.pop()
        #recomputeRange() returns true if the range changed
        if (insn.recomputeRange()):
            worklist.appendAll(insn.uses())
\end{verbatim}
}
\end{slide}

\begin{slide}
\begin{center}
{\Huge \bf{Results}}
\end{center}
\begin{itemize}
\item Instrumented IonMonkey to output percentage of adds and
  subtracts eliminated on a number of standard benchmark tests
\item Encouraging results
\end{itemize}
\centering
\includegraphics[width=0.98\textwidth]{counts_graph.png}
\end{slide}

\begin{slide}
\begin{center}
{\Huge \bf{Results }}
\end{center}
\begin{itemize}
\item Tested with three benchmark suites: v8bench, Sunspider, and Kraken
\item No statistically significant effect on v8bench
\item Tiny performance regression on Sunspider (0.1\%)
\item 1.5\% performance improvement on Kraken
\item Real overall wins, but not much.
\item Branch prediction is good these days, overflow checks wind up as
  noise in real programs
\item Microbenchmarks can see 30\% improvement
\end{itemize}
\end{slide}
